---
title: "Updated Bayesian Ordinal Analysis"
output: html_document
format:
  html: 
    fig-format: svg
date: "2025-04-21"
---

```{r setup, echo = FALSE, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(here)
library(tidyverse)
library(jtools)
library(sjPlot)
library(lme4)
library(lmerTest)
library(ordinal)
library(bayesplot)


td = read.csv(here("data", "ratings_tidy.csv"))

```

## **RQ 1:** Does learning explicitly about bilingualism impact novice-intermediate L2 learners’ attitudes towards bilingualism and language learning? 

This one I don't think we can answer with the data (no pre/post design)

## **RQ 2:** How do the attitudes towards bilingualism and language learning differ between students who have received explicit instruction about bilingualism and those who have not? 

We have evidence that those with instruction answered more positively for 3 of 4 question types.


#### Statistical Analysis 

An Bayesian Ordinal regression fit in R with `brms` with default priors. Rating was predicted as a function of survey version (1 or 2) and question type (4 levels), with random intercepts for prompt and participant. 

```{r, echo = FALSE, warning=FALSE, message=FALSE}
td$version = as.factor(td$version)

ord_mod_all = brms::brm(rating ~ version*question_type + (1 | prompt) + 
                       (1 | participant_number),
                     family = "cumulative",
                     data = td, 
                     file = here("data", "models", "ord_mod_all.rds"))
```

#### Results 

See the figures below:

Figure 1 shows the distributions of ratings in all 4 question types.
Figure 2 is the posterior distribution of the ordinal model - the distributions represent plausible changes in log-odds for each question type and group. The negative estimate suggest that version 2 was more likely to have less favorable ratings for the most interested question [Median = -0.69; 95% HDI -1.03 - -0.33 ], semi-affected questions [Median = -0.52; 95% HDI -0.81 - -0.25 ] and the least affected questions [Median = -0.64; 95% HDI -0.97 - -0.30].
There was no evidence that the directly affected questions were rated distinctly by the two groups [Median = 0.20; 95% HDI -0.11 - 0.52].


```{r, fig_1, echo = FALSE, warning=FALSE, message=FALSE}
rating_ver = td %>% 
  group_by(version, question_type) %>% 
  summarise(n_t = n())

pct_df_preq_all = td %>%
  group_by(version, rating, question_type) %>% 
  summarise(n = n()) %>% 
  left_join(rating_ver) %>% 
  mutate(pct = n/n_t)

# Modify the data
pct_df_preq_all$question_type <- factor(pct_df_preq_all$question_type, levels = c("directly_affected_questions", "least_affected_questions", "most_interested_qs", "semi_affected_questions"), 
                  labels = c("Directly affected questions", "Least affected questions", "Most important questions", "Semi-affected questions"))

pct_df_preq_all$question_type_f = factor(pct_df_preq_all$question_type, levels=c('Most important questions', 'Directly affected questions','Semi-affected questions','Least affected questions'))


pct_df_preq_all %>% 
  filter(!is.na(question_type)) %>% 
  ggplot(aes(y = pct, x = rating, fill = as.factor(version))) + 
  geom_col(color = "black", position = "dodge") + 
  scale_fill_manual(values = c("#71797E", "seagreen")) +
  facet_grid(question_type_f~version) + theme_apa() + xlab("Rating choice") +
  ylab("Percentage of Ratings") + 
  theme(strip.text.y = element_text(angle = 0),
        legend.position = "none")
```

```{r, fig_2, echo = FALSE, warning=FALSE, message=FALSE}
posterior <- as.matrix(ord_mod_all)

plot_title <- ggtitle("Posterior distributions",
                      "with medians and 80% intervals")
mcmc_areas(posterior,
           pars = c("b_version2", "b_version2:question_typeleast_affected_questions", "b_version2:question_typemost_interested_qs", "b_version2:question_typesemi_affected_questions"),
           prob = 0.8) + plot_title +
  scale_y_discrete(limits = c("b_version2:question_typeleast_affected_questions",
                                       "b_version2:question_typesemi_affected_questions",
                                       "b_version2", "b_version2:question_typemost_interested_qs")) + theme_minimal()

```


## **RQ 3:** Does learning explicitly about bilingualism impact students differently based on proficiency level?

It does not appear so - there is no evidence that class level impacted the size of the effect of having explicit instruction.


#### Statistical Analysis 

An Bayesian Ordinal regression fit in R with `brms` with default priors. Rating was predicted as a function of survey version (1 or 2) and current class (4 levels), with random intercepts for prompt and participant. 

```{r, warning = FALSE, message=FALSE}

td$current_class = as.factor(td$current_class)

ord_mod_class = brms::brm(rating ~ current_class*version + (1 | prompt) + 
                          (1 | participant_number),
                        family = "cumulative",
                        data = td, 
                        file = here("data", "models", "ord_mod_class.rds"))
```

```{r, fig_3, echo = FALSE, warning = FALSE, message=FALSE}
rating_ver_class = td %>% 
  group_by(version, current_class) %>% 
  summarise(n_t = n())

pct_df_preq_all_class = td %>%
  group_by(version, rating) %>% 
  summarise(n = n()) %>% 
  left_join(rating_ver_class) %>% 
  mutate(pct = n/n_t)

# Modify the data

pct_df_preq_all_class %>% 
  ggplot(aes(y = pct, x = rating, fill = as.factor(version))) + 
  geom_col(color = "black", position = "dodge") + 
  scale_fill_manual(values = c("#71797E", "seagreen")) +
  facet_grid(current_class~version) + theme_apa() + xlab("Rating choice") +
  ylab("Current Class") + 
  theme(strip.text.y = element_text(angle = 0),
        legend.position = "none")
```


```{r, fig_4, echo = FALSE, warning = FALSE, message=FALSE}
posterior_o<- as.matrix(ord_mod_class)

plot_title <- ggtitle("Posterior distributions",
                      "with medians and 80% intervals")
mcmc_areas(posterior_o,
           pars = c("b_version", 
                    "b_current_class2:version", "b_current_class3:version", "b_current_class4:version"),
           prob = 0.8) + plot_title + theme_minimal()

```


