---
title: "Survey initial analysis"
output: html_document
date: "2025-02-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(here)
library(tidyverse)
library(jtools)

td = read.csv(here("data", "ratings_tidy.csv"))
```

<!-- @tbl-log-model-->

<!---```{r}
#| label: tbl-log-model
#| tbl-cap: "Summary of the posterior distribution modeling of the probability of guessing an emotion adjective as a function of animacy and frame The table includes posterior medians, the 95% HDI, the percentage of the HDI within the ROPE, and the maximum probability of effect (MPE)."

read_csv(here("docs", "tables", "study_2_model.csv"), col_types = cols(.default = "c")) %>% 
  knitr::kable(format = "pandoc", align = c("l", "r", "r", "r", "r", "r"),
    label = "study2-mod-log")
```
-->

# Analysis 

The analysis is broken into sections. The first concerns the most important 5 questions that should be directly affected by the lesson, followed by the remaining directly affected questions, semi-affected questions and ending with the least affected questions. 
The plots show the percentage of each rating for each question type (instead of total count since groups were unequal).


## Most important 5 questions:

1. How much do you like the idea of being/becoming bilingual? `like_becoming_bilingual` 

2. Even if the other person knows both languages, how rude is it to switch between languages in a conversation? (reverse coded) `codeswitch_rude`

3. How important is Spanish in the United States? `important_spanish`

4. How valuable is it to be bilingual? `valuable`

5. How important is it to not have an accent when speaking a second language? (reverse code) `no_accent`


```{r, include=FALSE}
rating_ver = td %>% 
  filter(question_type == "most_interested_qs") %>% 
  group_by(version) %>% 
  summarise(n_t = n())


pct_df_preq = td %>% 
  filter(question_type == "most_interested_qs") %>% 
  group_by(version, rating) %>% 
  summarise(n = n()) %>% 
  left_join(rating_ver) %>% 
  mutate(pct = n/n_t)
```

```{r}
#| label: pct1
#| fig-cap: "The percentage of each rating by group for the 5 questions in focus"
pct_df_preq %>% 
  ggplot(aes(y = pct, x = rating, fill = as.factor(version))) + 
  geom_col(color = "black", position = "dodge") + 
  scale_fill_manual(values = c("#71797E", "seagreen")) +
  facet_wrap(~version) + theme_apa() + xlab("Rating choice") +
  ylab("Percentage of Ratings")
```

<!--- The figure shows the percentage of each rating for the groups 1 and 2.
Overall, group 1 rated the most five important questions as 5 more often than group 2 (`r round(pct_df_preq$pct[5], digits = 2)*100` percent and `r round(pct_df_preq$pct[10], digits = 2)*100` percent respectively).
Group 1 chose 4 `r round(pct_df_preq$pct[4], digits = 2)*100` percent of the time, which was slightly less than group 2 `r round(pct_df_preq$pct[10], digits = 2)*100`.
Fewer than 14 percent of responses were ratings of 1, 2, or 3 in both groups. -->

## The rest of the directly affected questions:

1. Should all schools in the United States teach children to speak in more than one language? `schools_teach` 

2. Is learning Spanish worth the effort it takes?
`effort`

3. Should people learn another language even if they don't plan on traveling abroad?
`learn_not_abroad` **I don't see this in the data**

4. Is speaking or learning another language important to you because it will help you in your job or future career?
`future_career`

5. Is speaking or learning another language important to you because it will help you connect with people who speak that language?
`connect_with_people`

6. Are multilingual people an asset to the United States?
`multilingualism_asset`

7. Is it acceptable for people to be monolingual in today's world? (reverse code)
`acceptable_monolingualism` **I don't see this in the data**

8. Do you think you can become bilingual?
become_bilingual `become_bilingual`


```{r, include=FALSE}
rating_ver_da = td %>% 
  filter(question_type == "directly_affected_questions") %>% 
  group_by(version) %>% 
  summarise(n_t = n())

pct_df_preq_da = td %>% 
  filter(question_type == "directly_affected_questions") %>% 
  group_by(version, rating) %>% 
  summarise(n = n()) %>% 
  left_join(rating_ver_da) %>% 
  mutate(pct = n/n_t)
```

```{r}
#| label: pct3
#| fig-cap: "The percentage of each rating by group for the directly affected questions"
pct_df_preq_da %>% 
  ggplot(aes(y = pct, x = rating, fill = as.factor(version))) + 
  geom_col(color = "black", position = "dodge") + 
  scale_fill_manual(values = c("#71797E", "seagreen")) +
  facet_wrap(~version) + theme_apa() + xlab("Rating choice") +
  ylab("Percentage of Ratings")
```

## Semi-affected questions:

1. How much smarter does knowing two languages make a person?
`smarter_two_languages`

2. How much do you wish you started learning Spanish at a younger age?
`started_younger`

3. How much do you admire people who speak multiple languages?
`admire_multilinguals`

4. How much do you like learning languages?
`like_langauges`

5. How much interest do you have in learning a language?
`interest_languages`

6. How much do you like Spanish?
`like_spanish`

7. How much do you like hearing Spanish?
`like_hearing_spanish`

8. How much do you like communicating with Spanish speakers?
`communicating_spanish_speakers`

9. Do people who speak more than one language have a wider world view than monolinguals?
`worldview`

10. How boring are Spanish classes? (reverse coded)
`boring_spanish_class`

11. How fun is studying languages?
`fun_studying_languages`

12. How interesting is studying languages?
`interesting_studying_languages`

13. How interesting do you find Spanish-speaking cultures?
`interesting_spanish_cultures`

14. How beautiful of a language is Spanish?
`beautiful`

```{r, include=FALSE}
rating_ver_sa = td %>% 
  filter(question_type == "semi_affected_questions") %>% 
  group_by(version) %>% 
  summarise(n_t = n())

pct_df_preq_sa = td %>% 
  filter(question_type == "semi_affected_questions") %>% 
  group_by(version, rating) %>% 
  summarise(n = n()) %>% 
  left_join(rating_ver_sa) %>% 
  mutate(pct = n/n_t)
```

```{r}
#| fig-cap: "The percentage of each rating by group for the semi-affected questions"
pct_df_preq_sa %>% 
  ggplot(aes(y = pct, x = rating, fill = as.factor(version))) + 
  geom_col(color = "black", position = "dodge") + 
  scale_fill_manual(values = c("#71797E", "seagreen")) +
  facet_wrap(~version) + theme_apa() + xlab("Rating choice") +
  ylab("Percentage of Ratings")
```


## Least affected questions:

1. How often are you excited to go to Spanish class?
`excited_spanish_class`

2. How often do you seek out ways to engage with Spanish outside of Spanish class and homework?
`outside_engagement`

3. How often is Spanish one of your favorite classes?
`spanish_favorite_class`

4. How difficult is it to learn Spanish in a language classroom? (reverse coded)
`difficult_classroom_learning`



```{r, include=FALSE}
rating_ver_la = td %>% 
  filter(question_type == "least_affected_questions") %>% 
  group_by(version) %>% 
  summarise(n_t = n())


pct_df_preq_la = td %>% 
  filter(question_type == "least_affected_questions") %>% 
  group_by(version, rating) %>% 
  summarise(n = n()) %>% 
  left_join(rating_ver) %>% 
  mutate(pct = n/n_t)

```

1. How often are you excited to go to Spanish class?
`excited_spanish_class`

2. How often do you seek out ways to engage with Spanish outside of Spanish class and homework?
`outside_engagement`

3. How often is Spanish one of your favorite classes?
`spanish_favorite_class`

4. How difficult is it to learn Spanish in a language classroom? (reverse coded)
`difficult_classroom_learning`

```{r}
#| label: pct2
#| fig-cap: "The percentage of each rating by group for the least effected questions"
pct_df_preq_la %>% 
  ggplot(aes(y = pct, x = rating, fill = as.factor(version))) + 
  geom_col(color = "black", position = "dodge") + 
  scale_fill_manual(values = c("#71797E", "seagreen")) +
  facet_wrap(~version) + theme_apa() + xlab("Rating choice") +
  ylab("Percentage of Ratings")

```

<!--- 
For both groups, 3 was the most common rating: `r round(pct_df_preq_la$pct[3], digits = 2)*100` percent for group 1 and 
`r round(pct_df_preq_la$pct[8], digits = 2)*100` percent for group 2. 
Both groups picked 5 less than 12 percent of the time (`r round(pct_df_preq_la$pct[5], digits = 2)*100` for group 1 and `r round(pct_df_preq_la$pct[10], digits = 2)*100` for group 2).
--->


